---
title: "p8105_hw3_cl3938"
author: "Cynthia Liu (cl3938)"
output: github_document
---

```{r setup}
library(tidyverse)
library(p8105.datasets)
library(dplyr)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

##Problem 1

###Load the data

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```


Let's make a plot

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


Let's make a table!!

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```
#Question 2

###Load, tidy, and wrangle data
``` {r}
accel = 
  read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(cols = starts_with('activity'), names_to = 'minute', names_prefix = "activity.", values_to = 'activity') %>%
  mutate(
    minute = as.numeric(minute),
    day = as.factor(day),
    activity = as.numeric(activity),
    day = forcats::fct_relevel(day, c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")),
    day_type = ifelse(day %in% c("Saturday", "Sunday"), "Weekend", "Weekday" )) %>%
  relocate(day_type, .after = day)
```

This dataset contains `r nrow(accel)` observations and `r ncol(accel)` variables. 

Observations are at the level of daily activity of the patient, with data organized per minute of the day. There is information on the week, as well as on the day (ID of the day, day of the week, weekend vs weekday), and the remaining columns are for activity by the minute, numbered by the chronological order of minutes. 

###Generate daily activity table

``` {r}
accel %>% 
  group_by(week, day_id, day) %>%
  summarize(total_activity = sum(activity)) %>%
	knitr::kable(caption = 'Total Activity per Day')
```
The activity data doesn't show any apparent trends, however, it is interesting to note that the patient had little to no activity during the last two Saturdays of this dataset, either due to a technology malfunction or just lower weekend movement.

###Create accelerometer plot

```{r}
accel %>%
  ggplot(aes(x = minute, y = activity, color = day)) +
  geom_line() +
  labs(
    title = "Accelerometer Activity over Time",
    x = "Minute of the Day",
    y = "Activity"
  )
```

#Question 3

```{r}
data("ny_noaa")
```

This dataset is NOAA data 

###Clean data

``` {r}
noaa_df <- ny_noaa %>%
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>%
  mutate(
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    prcp = prcp/10,
    tmax = tmax/10,
    tmin = tmin/10
    ) 
```

```{r}
noaa_df %>%
  count(snow) %>%
  arrange(desc(n))
```
The most commonly observed values for snowfall are 0 and NA, which make sense. The 0 value makes sense because not every location snows every winter, and snow only occurs every once in a while. The NA value also makes sense - due to the nature of the data, some locations only record a subset of variables, and it's possible locations are not recording snow or that they are not recording it during certain times of the year, like summer.

###Two-panel plot of average max temperatures

``` {r}
months.labs <- c("January", "July")
names(months.labs) <- as.numeric(c("1", "7"))

noaa_df %>%
  group_by(id, year, month) %>%
  filter(month == 1 || month == 7, na.rm = TRUE) %>%
  summarize (mean_tmax = mean(tmax)) %>%
  ggplot(aes(x=year, y = mean_tmax, group = id)) +
    geom_point(alpha = .5) +
    facet_grid(
      .~month, 
      labeller = labeller(month = months.labs)) +
    labs(
      title = "NY average max temperatures in January and July",
      x = "Year",
      y = "Mean max temp (C)",
      caption = "Data from NOAA (1981-2010) "
    ) +
    theme(panel.spacing = unit(2, "lines"))

```
In terms of observable structure, average max temperatures are high in July (hovering around 20-30 degrees Celsius), and low in January (around 0 degrees Celsius) across the entire time period. There are no significant trends in max temperature change over time. However, there are a few outliers: in January, very low temperatures at a station in 1982, and in July, a very low max temperature in 1988. 

###Two-panel plot of tmax vs tmin for full dataset

```{r}
  tmax_tmin_p = noaa_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
    geom_hex() +
    labs (
      title = "Max Temp by Min Temp",
      x = "Minimum Temperature (C)",
      y = "Maxmimum Temperature (C)",
      caption = "Data from NOAA (1981-2010)"
    ) +
    theme(legend.position = "right")
```


###Distribution of snowfall values > 0 and <100 by year 

```{r}

snow_year_p =
  noaa_df %>%
  filter(snow > 0 & snow < 100, na.rm = TRUE) %>%
  group_by(year) %>%
  ggplot(aes(x=snow, y = year, group = year)) +
    geom_density_ridges() +
    labs (
      title = "Distribution of Snowfall",
      x = "Snowfall (mm)",
      y = "Year",
      caption = "Data from NOAA"
    )
```

###Create final plot

```{r}
tmax_tmin_p + snow_year_p
```

